Monitoring the Evolution of Flask \& Python-Based Web Services

### 2018-04-09 discussion.tex
retitled

  \subsection*{Lack of Empirical Proof of Ease of Adoptability}

    We designed the dashboard to make it easy to adopt and we think this is to a certain degree apparent. However, we have only presented a single use case where the API developers have used the dashboard. It would be worthwhile to organize a dedicated study to observe the challenges and opportunities of the adoption of such a dashboard across different applications. It will also be important to see whether the perspectives presented here are relevant for other developers, and also which other perspectives are missing.

@@

>>>>

@@

  \subsection*{Need of Further Proof of Usefulness and Ease of Adoptability}

    We designed the dashboard to make it easy to adopt and we think this is to a certain degree apparent. However, we have only presented a single use case where the API developers have used the dashboard. It has been successfull, since based on the dashboard, the developers have become aware of the limitations of their service APIs and have started improving it. However, we belive that it will be worthwhile to organize a dedicated study to observe the challenges and opportunities of the adoption of such a dashboard across different applications. It will also be important to see whether the perspectives presented here are relevant for other developers, and also which other perspectives are missing.

### 2018-04-09 paper.tex
lay outing...
retitled

%
\title{Agile Monitoring of Evolving Web Services: For the Little Man}

@@

In the previous sections we discussed our proposal for a monitoring dashboard for Flask-based web applications as the means for providing a low effort and flexible analytics solution to projects on a budget. The emphasis is in allowing application developers to gain insight into how the performance and utilization of their services evolves together with the application itself. For this purpose the \tool allows for integration with both \git and the Travis continuous integration platform for a seamless user experience. We also discussed more advanced features of the dashboard in the form of grouping results by user groups, and provided evidence towards the use of unit testing-sourced synthetic loads as a limited form of performance prediction. Finally we measured the overhead of the proposed solution within acceptable by the case study application owner limits, and discussed what we perceive as limitations of our proposal in Section~\ref{sec:discussion}.

The same section effectively acts also as a roadmap for further development of the \tool. Some of the tasks have already been identified as ongoing work (allowing different types of grouping, enabling performance prediction, developing a meta-dashboard) while others constitute short- to medium-term future work (identification of important versions to be monitored, support for endpoint provenance) with the ultimate goal of increasing the adoption of the \tool by as many projects as possible. This will allow us to evaluate our proposal in depth in the future.

\va{@Mircea: see what you can add in the above, if any}

>>>>

% An Agile Analytics Platform for Visualizing Evolving Flask-Based Python Web Services
% A Dashboard for the Agile Monitoring of Evolving Web Services: For the Little Man
% A Dashboard for Monitoring Flask \& Python-Based Web Service Evolution
% Monitoring the Evolution of Flask \& Python-Based Web Services
% A Dashboard for Monitoring the Evolution of Flask \& Python-Based Web Services
\title{Monitoring the Evolution of Flask \& Python-Based Web Services}

@@

In the previous sections we discussed our proposal for a monitoring dashboard for Flask-based web applications as the means for providing a low effort and flexible analytics solution to projects on a budget. The emphasis is in allowing application developers to gain insight into how the performance and utilization of their services evolves together with the application itself. For this purpose the \tool allows for integration with both \git and the Travis CI platform. We also discussed more advanced features of the dashboard in the form of grouping results by user groups, and provided evidence towards the use of unit testing-sourced synthetic loads as a potential limited form of performance prediction. Finally we measured the overhead of the proposed solution within acceptable by the case study application owner limits. Section~\ref{sec:discussion} which discusses what we perceive as limitations of our proposal and also potential challenges for other similar tools, effectively acts also as a roadmap for further development, with the ultimate goal of increasing the adoption of the \tool by as many projects as possible. This will allow us to evaluate our proposal in depth in the future.

% Some of the tasks have already been identified as ongoing work (allowing different types of grouping, enabling performance prediction, developing a meta-dashboard) while others constitute short- to medium-term future work (identification of important versions to be monitored, support for endpoint provenance)

### 2018-04-09 related-work.tex
retitled

Service monitoring data are increasingly being used and more across a variety of systems in which adaptability, flexibility, and environmental awareness are essential~\cite{pernici2016monitoring}. Monitoring can be performed on different types of system components and for several purposes. Service-oriented literature on the subject discussed the multitude of aspects related to monitoring for services; the interested reader is referred to ~\cite{ghezzi2007run},~\cite{metzger2010analytical}, or more recently~\cite{pernici2016monitoring} for an extensive presentation of the subject.

In this context, our work falls within the server-side run-time monitoring of services ~\cite{ghezzi2007run}. While we don't implement the more advanced features of related monitoring solutions like QoS policies driving the monitoring, it presents nevertheless an easy to use approach towards improving the performance of web applications. The emphasis in this work is given in identifying the appropriate way to visualize the monitoring data.

@@

@@

>>>>

% testing... the performacen
% there is plenty of work on testing for finding fauls
% people: https://scholar.google.nl/citations?user=DWd972sAAAAJ&hl=en&oi=ao

A rich body of work exists that studies the evolution and maintenance of APIs from the perspective of source code evolution \cite{dig2006apis, hora2015developers, hora2015apiwave}. However, since services are increasingly being used across a variety of systems in which adaptability, flexibility, and environmental awareness are essential~\cite{pernici2016monitoring} in our work we focus on the analysis of the runtime aspects of service APIs. Our work is thus, more related to the work on dynamic analysis of software systems; however, most of this work has been done in order to support the reverse engineering of software systems\cite{Corn09-dynamic}.

In this context, our work falls within the server-side run-time monitoring of services ~\cite{ghezzi2007run}. While we don't implement the more advanced features of related monitoring solutions like QoS policies driving the monitoring, it presents nevertheless an easy to use approach towards improving the performance of web applications.

@@

Kieker -- has been presented by Rohr et al \cite{Rohr08} as a system targeted at monitoring the performance of Java services. The system provides a better performance in terms of overhead, but does not support integration with git and multi-version analysis as the system presented here.

@@

Monitoring can be performed on different types of system components and for several purposes beyond the ones presented here. Service-oriented literature on the subject discussed the multitude of aspects related to monitoring for services; the interested reader is referred to ~\cite{ghezzi2007run},~\cite{metzger2010analytical}, or more recently~\cite{pernici2016monitoring} for an extensive presentation of the subject.

% Our 
% - Lanza has done nice stuff for source code evolution patterns. we need more work on automatic detection of service / endpoint evolution patterns. as exemplified by the utilization evolution.

### 2018-04-09 testing.tex
retitled

  \begin{lstlisting}[style=custompython]  
# LOC #4: to be added to the .travis.yml file
python -m flask_monitoringdashboard.collect
   --test_folder=./tests_zeeguu_api --times=5 
   --url=https://zeeguu.unibe.ch/api/dashboard

@@

      \begin{figure}[h!]
        \centering
        \includegraphics[width=0.85\columnwidth]{travis_builds-bw}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}

>>>>

  \begin{lstlisting}[style=custompython]

# LOC #4: is added to the .travis.yml file
python -m flask_monitoringdashboard.collect
   --test_folder=./tests_zeeguu_api --times=5 
   --url=https://zeeguu.unibe.ch/api/dashboard

@@

      \begin{figure}[h!]
        \centering
        \includegraphics[width=\columnwidth]{travis_builds-bw}
        \caption{Response times in 5 consequent Travis builds}
        \label{fig:builds}
      \end{figure}
